# Control flow

## Exercise 5.2.4

**Q1.** `ifelse()` return type

It's type unstable, i.e. the type of return will depend on the type of each condition (`yes` and `no`, i.e.), and works only for cases where `test` argument evaluated to a `logical` type.

```{r}
ifelse(TRUE, 1, "no") # numeric returned
ifelse(FALSE, 1, "no") # character returned
ifelse(NA, 1, "no")
```

Additionally, if the test doesn't resolve to `logical` type, it will try to coerce whatever the resulting type to `logical`:

```{r}
# will work
ifelse("TRUE", 1, "no")
ifelse("true", 1, "no")

# won't work
ifelse("tRuE", 1, "no")
ifelse(NaN, 1, "no")
```

**Q2.** Why does the following code work?

As can be seen, the code works because the tests are successfully coerced to a logical type.

```{r}
x <- 1:10
as.logical(length(1:10))
if (length(x)) "not empty" else "empty"

x <- numeric()
as.logical(length(numeric()))
if (length(x)) "not empty" else "empty"
```

## Exercise 5.3.3

**Q1.** Why does this work?

This works because `1:length(x)` goes both ways; in this case, from 1 to 0. And, since out-of-bound values for atomic vectors is `NA`, all related operations with it also lead to `NA`.

```{r}
x <- numeric()
out <- vector("list", length(x))

for (i in 1:length(x)) {
  print(x[i])
  print(out[i])

  out[i] <- x[i]^2
}

out
```

A way to do avoid this unintended behavior would be:

```{r, error=TRUE}
x <- numeric()
out <- vector("list", length(x))

for (i in 1:seq_along(x)) {
  out[i] <- x[i]^2
}

out
```

**Q2.** Index vectors

Surprisingly (at least to me), `x` takes all values of the vector `xs`:

```{r}
# xs <- c(1, 2, 3)
xs <- c(4, 5, 6)
for (x in xs) {
  print(x)
  xs <- c(xs, x * 2)
}

xs
```

**Q3.** Index increase

In a `for` loop - like in a `while` loop - the index is updated in the beginning of each iteration.

```{r}
for (i in 1:3) {
  cat("before: ", i, "\n")
  i <- i * 2
  cat("after: ", i, "\n")
}

i <- 1
while (i < 3) {
  cat("before: ", i, "\n")
  i <- i * 2
  cat("after: ", i, "\n")
}
```
