# Control flow

```{r Control-flow-1, include = FALSE}
source("common.R")
```

## Choices (Exercises 5.2.4)

**Q1.** What type of vector does each of the following calls to `ifelse()` return?

```{r Control-flow-2, eval = FALSE}
ifelse(TRUE, 1, "no")
ifelse(FALSE, 1, "no")
ifelse(NA, 1, "no")
```

Read the documentation and write down the rules in your own words.

**A1.** Here are the rules about what a call to `ifelse()` might return: 

- It is type unstable, i.e. the type of return will depend on the type of which condition is true (`yes` or `no`, i.e.): 

```{r Control-flow-3}
ifelse(TRUE, 1, "no") # `numeric` returned
ifelse(FALSE, 1, "no") # `character` returned
```

- It works only for cases where `test` argument evaluates to a `logical` type:

```{r Control-flow-4}
ifelse(NA_real_, 1, "no")
ifelse(NaN, 1, "no")
```

- If `test` is argument is of logical type, but `NA`, it will return `NA`:

```{r Control-flow-5}
ifelse(NA, 1, "no")
```

- If the `test` argument doesn't resolve to `logical` type, it will try to coerce the output to a `logical` type:

```{r Control-flow-6}
# will work
ifelse("TRUE", 1, "no")
ifelse("false", 1, "no")

# won't work
ifelse("tRuE", 1, "no")
ifelse(NaN, 1, "no")
```

This is also clarified in the docs for this function:

> A vector of the same length and attributes (including dimensions and `"class"`) as `test` and data values from the values of `yes` or `no`. The mode of the answer will be coerced from logical to accommodate first any values taken from yes and then any values taken from `no`.

**Q2.** Why does the following code work?

```{r Control-flow-7}
x <- 1:10
if (length(x)) "not empty" else "empty"

x <- numeric()
if (length(x)) "not empty" else "empty"
```

**A2.** The code works because the conditional expressions in `if()` - even though of `numeric` type - can be successfully coerced to a `logical` type.

```{r Control-flow-8}
as.logical(length(1:10))

as.logical(length(numeric()))
```

## Loops (Exercises 5.3.3)

**Q1.** Why does this code succeed without errors or warnings? 
    
```{r Control-flow-9, results = FALSE}
x <- numeric()
out <- vector("list", length(x))
for (i in 1:length(x)) {
  out[i] <- x[i]^2
}
out
```

**A1.** This works because `1:length(x)` works in both positive and negative directions.

```{r Control-flow-10}
1:2
1:0
1:-3
```

In this case, since `x` is of length `0`, `i` will go from `1` to `0`. 

Additionally, since out-of-bound (OOB) value for atomic vectors is `NA`, all related operations with OOB values will also produce `NA`.

```{r Control-flow-11}
x <- numeric()
out <- vector("list", length(x))

for (i in 1:length(x)) {
  print(paste("i:", i, ", x[i]:", x[i], ", out[i]:", out[i]))

  out[i] <- x[i]^2
}

out
```

A way to do avoid this unintended behavior is to use `seq_along()` instead:

```{r Control-flow-12}
x <- numeric()
out <- vector("list", length(x))

for (i in seq_along(x)) {
  out[i] <- x[i]^2
}

out
```

**Q2.** When the following code is evaluated, what can you say about the vector being iterated?

```{r Control-flow-13}
xs <- c(1, 2, 3)
for (x in xs) {
  xs <- c(xs, x * 2)
}
xs
```

**A2.** The iterator variable `x` initially takes all values of the vector `xs`. We can check this by printing `x` for each iteration:

```{r Control-flow-14}
xs <- c(1, 2, 3)
for (x in xs) {
  cat("x:", x, "\n")
  xs <- c(xs, x * 2)
  cat("xs:", paste(xs), "\n")
}
```

It is worth noting that `x` is not updated *after* each iteration; otherwise, it will take increasingly bigger values of `xs`, and the loop will never end executing.

**Q3.** What does the following code tell you about when the index is updated?

```{r Control-flow-15}
for (i in 1:3) {
  i <- i * 2
  print(i)
}
```

**A3.** In a `for()` loop the index is updated in the **beginning** of each iteration. Otherwise, we will encounter an infinite loop.

```{r Control-flow-16}
for (i in 1:3) {
  cat("before: ", i, "\n")
  i <- i * 2
  cat("after:  ", i, "\n")
}
```

Also, worth contrasting the behavior of `for()` loop with that of `while()` loop:

```{r Control-flow-17}
i <- 1
while (i < 4) {
  cat("before: ", i, "\n")
  i <- i * 2
  cat("after:  ", i, "\n")
}
```

## Session information

```{r Control-flow-18}
sessioninfo::session_info(include_base = TRUE)
```
