# Functionals

## Exercise 9.2.6 {-}

### Q1. Study `as_mapper()` {-}

```{r}
library(purrr)

# mapping by position -----------------------

x <- list(1, list(2, 3, list(1, 2)))

map(x, 1)
as_mapper(1)

map(x, list(2, 1))
as_mapper(list(2, 1))

# mapping by name -----------------------

y <- list(
  list(m = "a", list(1, m = "mo")),
  list(n = "b", list(2, n = "no"))
)

map(y, "m")
as_mapper("m")

# mixing position and name
map(y, list(2, "m"))
as_mapper(list(2, "m"))

# compact functions ----------------------------

map(y, ~ length(.x))
as_mapper(~ length(.x))
```

You can extract attributes using `purrr::attr_getter()`:

```{r}
pluck(Titanic, attr_getter("class"))
```

### Q2. Properly specifying anonymous functions {-}

As shown by `as_mapper()` outputs below, the second call is not appropriate for generating random numbers because it translates to `pluck()` function where the indices for plucking are taken to be randomly generated numbers.

```{r}
library(purrr)

map(1:3, ~ runif(2))
as_mapper(~ runif(2))

map(1:3, runif(2))
as_mapper(runif(2))
```

### Q3. Use the appropriate `map()` function {-}

Compute the standard deviation of every column in a numeric data frame.

```{r}
map_dbl(mtcars, sd)
```

Compute the standard deviation of every numeric column in a mixed data frame. 

```{r}
keep(iris, is.numeric) %>%
  map_dbl(sd)
```

Compute the number of levels for every factor in a data frame.

```{r}
modify_if(dplyr::starwars, is.character, as.factor) %>%
  keep(is.factor) %>%
  map_int(~ length(levels(.)))
```

### Q4. Statistics and visualization with `{purrr}` {-}

Extract the *p*-value from each test, then visualise.

```{r}
trials <- map(1:100, ~ t.test(rpois(10, 10), rpois(7, 10)))

p <- map_dbl(trials, "p.value")

plot(p)
```

### Q5. Fixing non-functioning code {-}

```{r}
x <- list(
  list(1, c(3, 9)),
  list(c(3, 6), 7, c(4, 7, 6))
)

triple <- function(x) x * 3
map(x, .f = ~ map(., ~ triple(.)))
```

### Q6. Use `map()` to fit linear models to the `mtcars` dataset {-}

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

map(formulas, ~ lm(formula = ., data = mtcars))
```

### Q7. Computing R-squared {-}

```{r}
bootstrap <- function(df) {
  df[sample(nrow(df), replace = TRUE), , drop = FALSE]
}

bootstraps <- map(1:10, ~ bootstrap(mtcars))

map_dbl(
  bootstraps,
  ~ summary(lm(formula = mpg ~ disp, data = .))$r.squared
)
```

## Exercise 9.4.6 {-}

### Q1. Explain the results {-}

`modify()` returns the object of type same as the input. Since the input here is a dataframe of certain dimensions and `.f = 1` translates to plucking the first element in each column, it returns a dataframes with the same dimensions with the plucked element recycled across rows.

```{r}
head(modify(mtcars, 1))
```

### Q2. Use `iwalk()` instead of `walk2()` {-}

```{r, eval=FALSE}
# with walk2() -----------------------

cyls <- split(mtcars, mtcars$cyl)
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
walk2(.x = cyls, .y = paths, .f = write.csv)

# with iwalk --------------------------

cyls <- split(mtcars, mtcars$cyl)
names(cyls) <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
iwalk(cyls, ~ write.csv(.x, .y))
```

