# Functionals

## Exercise 9.2.6

Q1. Study `as_mapper()`

```{r}
library(purrr)

# mapping by position -----------------------

x <- list(1, list(2, 3, list(1, 2)))

map(x, 1)
as_mapper(1)

map(x, list(2, 1))
as_mapper(list(2, 1))

# mapping by name -----------------------

y <- list(
  list(m = "a", list(1, m = "mo")),
  list(n = "b", list(2, n = "no"))
)

map(y, "m")
as_mapper("m")

# mixing position and name
map(y, list(2, "m"))
as_mapper(list(2, "m"))

# compact functions ----------------------------

map(y, ~ length(.x))
as_mapper(~ length(.x))
```

You can extract attributes using `purrr::attr_getter()`:

```{r}
pluck(Titanic, attr_getter("class"))
```

Q2. Properly specifying anonymous functions

As shown by `as_mapper()` outputs below, the second call is not appropriate for generating random numbers because it translates to `pluck()` function where the indices for plucking are taken to be randomly generated numbers.

```{r}
library(purrr)

map(1:3, ~ runif(2))
as_mapper(~ runif(2))

map(1:3, runif(2))
as_mapper(runif(2))
```

Q3. Use the appropriate `map()` function

Compute the standard deviation of every column in a numeric data frame.

```{r}
map_dbl(mtcars, sd)
```

Compute the standard deviation of every numeric column in a mixed data frame. 

```{r}
keep(iris, is.numeric) %>%
  map_dbl(sd)
```

Compute the number of levels for every factor in a data frame.

```{r}
modify_if(dplyr::starwars, is.character, as.factor) %>%
  keep(is.factor) %>%
  map_int(~ length(levels(.)))
```

Q4. Statistics and visualization with `{purrr}`

Extract the *p*-value from each test, then visualise.

```{r}
trials <- map(1:100, ~ t.test(rpois(10, 10), rpois(7, 10)))

p <- map_dbl(trials, "p.value")

plot(p)
```

Q5. Fixing non-functioning code

```{r}
x <- list(
  list(1, c(3, 9)),
  list(c(3, 6), 7, c(4, 7, 6))
)

triple <- function(x) x * 3
map(x, .f = ~ map(., ~ triple(.)))
```

Q6. Use `map()` to fit linear models to the `mtcars` dataset

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

map(formulas, ~ lm(formula = ., data = mtcars))
```

Q7. Computing R-squared

```{r}
bootstrap <- function(df) {
  df[sample(nrow(df), replace = TRUE), , drop = FALSE]
}

bootstraps <- map(1:10, ~ bootstrap(mtcars))

map_dbl(
  bootstraps,
  ~ summary(lm(formula = mpg ~ disp, data = .))$r.squared
)
```

## Exercise 9.4.6

Q1. Explain the results

`modify()` returns the object of type same as the input. Since the input here is a dataframe of certain dimensions and `.f = 1` translates to plucking the first element in each column, it returns a dataframes with the same dimensions with the plucked element recycled across rows.

```{r}
head(modify(mtcars, 1))
```

Q2. Use `iwalk()` instead of `walk2()`

```{r, eval=FALSE}
# with walk2() -----------------------

cyls <- split(mtcars, mtcars$cyl)
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
walk2(.x = cyls, .y = paths, .f = write.csv)

# with iwalk --------------------------

cyls <- split(mtcars, mtcars$cyl)
names(cyls) <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
iwalk(cyls, ~ write.csv(.x, .y))
```

Q3. Explain the code

`map2()` supplies the functions defined in `.x = trans` as `f` in the anonymous functions, while the names of the columns defined in  `.y = mtcars[nm]` are picked up by `var` in the anonymous function. Note that the function is iterating over indices for vectors of transformations and column names.

```{r, eval=FALSE}
trans <- list(
  disp = function(x) x * 0.0163871,
  am = function(x) factor(x, labels = c("auto", "manual"))
)

nm <- names(trans)
mtcars[nm] <- map2(trans, mtcars[nm], function(f, var) f(var))
```

In the `map` approach, the function is iterating over indices for vectors of column names.

```{r, eval=FALSE}
mtcars[nm] <- map(nm, ~ trans[[.x]](mtcars[[.x]]))
```

Q4. Difference between `map2()` and `walk2()`

If we use `map2()`, it will work, but it will print `NULL` to the terminal for every element of the list.

```{r, eval=FALSE}
bods <- split(BOD, BOD$Time)
nm <- names(bods)
map2(bods, nm, write.csv)
```

## Exercise 9.6.3

Q1. Predicate functions

> A predicate is a function that returns a **single** `TRUE` or `FALSE`.

The `is.na()` function does not return a single value, but instead returns a vector and thus isn't a predicate function.

```{r}
# contrast the following behavior of predicate functions
is.character(c("x", 2))
is.null(c(3, NULL))

# with this behavior
is.na(c(NA, 1))
```

The closest equivalent of a predicate function in base-R is `anyNA()` function.

```{r}
anyNA(c(NA, 1))
```

Q2. Fix `simple_reduce`

Supplied function:

```{r}
simple_reduce <- function(x, f) {
  out <- x[[1]]
  for (i in seq(2, length(x))) {
    out <- f(out, x[[i]])
  }
  out
}
```

Struggles with inputs of length 0 and 1 because function tries to access out-of-bound values.

```{r, error=TRUE}
simple_reduce(numeric(), sum)
simple_reduce(1, sum)
simple_reduce(1:3, sum)
```

This problem can be solved by adding `init` argument, which supplies the default or initial value for the function to operate on:

```{r}
simple_reduce2 <- function(x, f, init = 0) {
  # initializer will become the first value
  if (length(x) == 0L) {
    return(init)
  }
  if (length(x) == 1L) {
    return(x[[1L]])
  }

  out <- x[[1]]

  for (i in seq(2, length(x))) {
    out <- f(out, x[[i]])
  }
  out
}
```

Let's try it out:

```{r}
simple_reduce2(numeric(), sum)
simple_reduce2(1, sum)
simple_reduce2(1:3, sum)
```

With a different kind of function:

```{r}
simple_reduce2(numeric(), `*`, init = 1)
simple_reduce2(1, `*`, init = 1)
simple_reduce2(1:3, `*`, init = 1)
```

And another one:

```{r}
simple_reduce2(numeric(), `%/%`)
simple_reduce2(1, `%/%`)
simple_reduce2(1:3, `%/%`)
```


## Exercise 9.7.3

Q1. 

Q2. `eapply()` and `rapply()`

> `eapply()` applies FUN to the named values from an environment and returns the results as a list.

> `rapply()` is a recursive version of lapply with flexibility in how the result is structured (how = "..").
