# Names and values

## 2.2.2 Exercises 

**Q1.** Explain the relationship

```{r}
a <- 1:10
b <- a
c <- b
d <- 1:10
```

All of these variable names are actively bound to the same value.

```{r}
library(lobstr)

obj_addr(a)
obj_addr(b)
obj_addr(c)
obj_addr(d)
```

**Q2.** Function object address

Following code verifies that indeed these calls all point to the same underlying function object.

```{r}
obj_addr(mean)
obj_addr(base::mean)
obj_addr(get("mean"))
obj_addr(evalq(mean))
obj_addr(match.fun("mean"))
```

**Q3.** Converting non-syntactic names 

The conversion of non-syntactic names to syntactic ones can sometimes corrupt the data. Some datasets may require non-syntactic names.

To suppress this behavior, one can set `check.names = FALSE`.

**Q4.** Behavior of `make.names()` 

It just prepends `X` in non-syntactic names and invalid characters (like `@`) are translated to `.`.

```{r}
make.names(c("123abc", "@me", "_yu", "  gh", "else"))
```

**Q5.** Why is `.123e1` not a syntactic name? 

Because it is parsed as a number.

```{r}
.123e1 < 1
```

## 2.3.6 Exercises 

**Q1.** Usefulness of `tracemem()`

`tracemem()` traces copying of objects in R, but since the object created here is not assigned a name, there is nothing to trace. 

```{r}
tracemem(1:10)
```

**Q2.** Why two copies when you run this code?

Were it not for `4` being a double - and not an integer (`4L`) - this would have been modified in place.

```{r}
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4
```

Try with integer:

```{r}
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4L
```

As for why this still produces a copy, this is from Solutions manual:

> Please be aware that running this code in RStudio will result in additional copies because of the reference from the environment pane.

**Q3.** Study relationship

```{r}
a <- 1:10
b <- list(a, a)
c <- list(b, a, 1:10)

ref(a)

ref(b)

ref(c)
```

**Q4.** List inside another list

```{r}
x <- list(1:10)
x
obj_addr(x)

x[[2]] <- x
x
obj_addr(x)

ref(x)
```

Figure here:
<https://advanced-r-solutions.rbind.io/images/names_values/copy_on_modify_fig2.png>

## 2.4.1 Exercises 

**Q1.** Object size difference between `{base}` and `{lobstr}`

> This function...does not detect if elements of a list are shared.

```{r}
y <- rep(list(runif(1e4)), 100)

object.size(y)

obj_size(y)
```

**Q2.** Misleading object size

These functions are not externally created objects in R, but are always available, so doesn't make much sense to measure their size.

```{r}
funs <- list(mean, sd, var)
obj_size(funs)
```

Nevertheless, it's still interesting that the addition is not the same as size of list of those objects.

```{r}
obj_size(mean)
obj_size(sd)
obj_size(var)

obj_size(mean) + obj_size(sd) + obj_size(var)
```

**Q3.** Predict object sizes

```{r}
a <- runif(1e6)
obj_size(a)

b <- list(a, a)
obj_size(b)
obj_size(a, b)

b[[1]][[1]] <- 10
obj_size(b)
obj_size(a, b)

b[[2]][[1]] <- 10
obj_size(b)
obj_size(a, b)
```

## 2.5.3 Exercises

**Q1.** Why not a circular list?

Copy-on-modify prevents the creation of a circular list.

```{r}
x <- list()

obj_addr(x)

tracemem(x)

x[[1]] <- x

obj_addr(x[[1]])
```

**Q2.** Performance of `for` loops with and without creating copies

Let's first microbenchmark functions that does and does not create copies for varying lengths of number of columns.

```{r}
library(bench)
library(tidyverse)

generateDataFrame <- function(ncol) {
  as.data.frame(matrix(runif(100 * ncol), nrow = 100))
}

withCopy <- function(ncol) {
  x <- generateDataFrame(ncol)
  medians <- vapply(x, median, numeric(1))

  for (i in seq_along(medians)) {
    x[[i]] <- x[[i]] - medians[[i]]
  }

  return(x)
}

withoutCopy <- function(ncol) {
  x <- generateDataFrame(ncol)
  medians <- vapply(x, median, numeric(1))

  y <- as.list(x)

  for (i in seq_along(medians)) {
    y[[i]] <- y[[i]] - medians[[i]]
  }

  return(y)
}

benchComparison <- function(ncol) {
  bench::mark(
    withCopy(ncol),
    withoutCopy(ncol),
    iterations = 100,
    check = FALSE
  ) %>%
    dplyr::select(expression:total_time)
}

nColList <- list(1, 10, 50, 100, 250, 500, 1000)

names(nColList) <- as.character(nColList)

benchDf <- purrr::map_dfr(
  .x = nColList,
  .f = benchComparison,
  .id = "nColumns"
)
```

Plotting these benchmarks reveals how the performance gets increasingly worse as the number of dataframes increases:

```{r}
ggplot(
  benchDf,
  aes(
    x = as.numeric(nColumns),
    y = median,
    group = as.character(expression),
    color = as.character(expression)
  )
) +
  geom_line() +
  labs(
    x = "Number of Columns",
    y = "Median Execution Time (ms)",
    colour = "Type of function"
  )
```


**Q3.** `tracemem()` on an environment

It doesn't work and the documentation makes it clear as to why:

> It is not useful to trace NULL, environments, promises, weak references, or external pointer objects, as these are not duplicated

```{r, error=TRUE}
e <- rlang::env(a = 1, b = "3")
tracemem(e)
```
