# Measuring performance

```{r, include = FALSE}
source("common.R")
```

## Exercise 23.2.4

**Q1.** Profiling function with `torture = TRUE`

Let's first source the functions mentioned in exercises.

```{r, warning=FALSE}
library(profvis)

source("profiling-exercises.R")
```

First, we try without `torture = TRUE`: it returns no meaningful results. 

```{r, error=TRUE}
profvis(f())
```

Maybe because the function runs too fast?

```{r}
bench::mark(f(), check = FALSE, iterations = 1000)
```

As mentioned in the docs, setting `torture = TRUE`

> Triggers garbage collection after every torture memory allocation call.

This process somehow never seems to finish and crashes the RStudio session when it stops!

```{r, eval = FALSE}
profvis(f(), torture = TRUE)
```

The question says that documentation for `rm()` may provide clues:

```{r}
rm
```

I still couldn't figure out why. I would recommend checking out the [official answer](https://advanced-r-solutions.rbind.io/measuring-performance.html#profiling).

## Exercise 23.3.3

**Q1.** Differences between `system.time()` and `bench::mark()`

```{r}
library(dplyr)

n <- 1e6
x <- runif(100)

# bench -------------------

bench_df <- bench::mark(
  sqrt(x),
  x^0.5,
  iterations = n
)

t_bench_df <- bench_df %>%
  dplyr::select(expression, time) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(mean = mean(unlist(time))) %>%
  dplyr::ungroup() %>%
  dplyr::select(-time)

# system.time -------------------

# garbage collection performed immediately before the timing
t1_systime_gc <- system.time(for (i in 1:n) sqrt(x), gcFirst = TRUE) / n
t2_systime_gc <- system.time(for (i in 1:n) x^0.5, gcFirst = TRUE) / n

# garbage collection not performed immediately before the timing
t1_systime_nogc <- system.time(for (i in 1:n) sqrt(x), gcFirst = FALSE) / n
t2_systime_nogc <- system.time(for (i in 1:n) x^0.5, gcFirst = FALSE) / n

t_systime_df <- tibble(
  "expression" = bench_df$expression,
  "systime_with_gc_us" = c(t1_systime_gc["elapsed"], t2_systime_gc["elapsed"]),
  "systime_with_nogc_us" = c(t1_systime_nogc["elapsed"], t2_systime_nogc["elapsed"])
) %>%
  dplyr::mutate(
    systime_with_gc_us = systime_with_gc_us * 1e6,
    systime_with_nogc_us = systime_with_nogc_us * 1e6
  )
```

Compare results from these alternatives:

```{r}
t_bench_df

t_systime_df
```

The comparison reveals that these two approaches yield quite similar results.

**Q2.** Microbenchmarking ways to compute square root

Microbenchmarking all ways to compute square root of a vector mentioned in this chapter.

```{r}
x <- runif(1000)

bench::mark(
  sqrt(x),
  x^0.5,
  x^(1 / 2),
  exp(log(x) / 2),
  iterations = 1000
) %>%
  dplyr::arrange(median)
```

The specialized primitive function `sqrt()` (written in `C`) is the fastest way to compute square root.
