# Measuring performance

```{r Perf-measure-1, include = FALSE}
source("common.R")
```

Attaching the needed libraries:

```{r Perf-measure-2, warning=FALSE, message=FALSE}
library(profvis, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
```

## Profiling (Exercises 23.2.4)

---

**Q1.** Profile the following function with `torture = TRUE`. What is surprising? Read the source code of `rm()` to figure out what's going on.

```{r Perf-measure-3, eval = FALSE}
f <- function(n = 1e5) {
  x <- rep(1, n)
  rm(x)
}
```

**A1.** Let's first source the functions mentioned in exercises.

```{r Perf-measure-4, warning=FALSE}
source("profiling-exercises.R")
```

First, we try without `torture = TRUE`: it returns no meaningful results. 

```{r Perf-measure-5, error=TRUE}
profvis(f())
```

Maybe because the function runs too fast?

```{r Perf-measure-6}
bench::mark(f(), check = FALSE, iterations = 1000)
```

As mentioned in the docs, setting `torture = TRUE`

> Triggers garbage collection after every torture memory allocation call.

This process somehow never seems to finish and crashes the RStudio session when it stops!

```{r Perf-measure-7, eval = FALSE}
profvis(f(), torture = TRUE)
```

The question says that documentation for `rm()` may provide clues:

```{r Perf-measure-8}
rm
```

I still couldn't figure out why. I would recommend checking out the [official answer](https://advanced-r-solutions.rbind.io/measuring-performance.html#profiling).

---

## Microbenchmarking (Exercises 23.3.3)

---

**Q1.** Instead of using `bench::mark()`, you could use the built-in function `system.time()`. But `system.time()` is much less precise, so you'll need to repeat each operation many times with a loop, and then divide to find the average time of each operation, as in the code below.

```{r Perf-measure-9, eval = FALSE}
n <- 1e6
system.time(for (i in 1:n) sqrt(x)) / n
system.time(for (i in 1:n) x^0.5) / n
```

How do the estimates from `system.time()` compare to those from `bench::mark()`? Why are they different?

**A1.** Let's benchmark first:

```{r Perf-measure-10}
n <- 1e6
x <- runif(100)

# bench -------------------

bench_df <- bench::mark(
  sqrt(x),
  x^0.5,
  iterations = n
)

t_bench_df <- bench_df %>%
  dplyr::select(expression, time) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(mean = mean(unlist(time))) %>%
  dplyr::ungroup() %>%
  dplyr::select(-time)

# system.time -------------------

# garbage collection performed immediately before the timing
t1_systime_gc <- system.time(for (i in 1:n) sqrt(x), gcFirst = TRUE) / n
t2_systime_gc <- system.time(for (i in 1:n) x^0.5, gcFirst = TRUE) / n

# garbage collection not performed immediately before the timing
t1_systime_nogc <- system.time(for (i in 1:n) sqrt(x), gcFirst = FALSE) / n
t2_systime_nogc <- system.time(for (i in 1:n) x^0.5, gcFirst = FALSE) / n

t_systime_df <- tibble(
  "expression" = bench_df$expression,
  "systime_with_gc_us" = c(t1_systime_gc["elapsed"], t2_systime_gc["elapsed"]),
  "systime_with_nogc_us" = c(t1_systime_nogc["elapsed"], t2_systime_nogc["elapsed"])
) %>%
  dplyr::mutate(
    systime_with_gc_us = systime_with_gc_us * 1e6,
    systime_with_nogc_us = systime_with_nogc_us * 1e6
  )
```

Compare results from these alternatives:

```{r Perf-measure-11}
t_bench_df

t_systime_df
```

The comparison reveals that these two approaches yield quite similar results.

---

**Q2.** Here are two other ways to compute the square root of a vector. Which do you think will be fastest? Which will be slowest? Use microbenchmarking to test your answers.

```{r Perf-measure-12, eval = FALSE}
x^(1 / 2)
exp(log(x) / 2)
```

---

**A2.** Microbenchmarking all ways to compute square root of a vector mentioned in this chapter.

```{r Perf-measure-13}
x <- runif(1000)

bench::mark(
  sqrt(x),
  x^0.5,
  x^(1 / 2),
  exp(log(x) / 2),
  iterations = 1000
) %>%
  dplyr::arrange(median)
```

The specialized primitive function `sqrt()` (written in `C`) is the fastest way to compute square root.

---

## Session information

```{r Perf-measure-14}
sessioninfo::session_info(include_base = TRUE)
```
