# Rewriting R code in C++

## Exercise 25.2.6

### Q1. Figure out base function corresponding to Rccp code {-}

```{r}
library(Rcpp)
```


```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double f1(NumericVector x) {
  int n = x.size();
  double y = 0;

  for(int i = 0; i < n; ++i) {
    y += x[i] / n;
  }
  return y;
}

// [[Rcpp::export]]
NumericVector f2(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] + x[i];
  }
  return out;
}

// [[Rcpp::export]]
bool f3(LogicalVector x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    if (x[i]) return true;
  }
  return false;
}

// [[Rcpp::export]]
int f4(Function pred, List x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    LogicalVector res = pred(x[i]);
    if (res[0]) return i + 1;
  }
  return 0;
}

// [[Rcpp::export]]
NumericVector f5(NumericVector x, NumericVector y) {
  int n = std::max(x.size(), y.size());
  NumericVector x1 = rep_len(x, n);
  NumericVector y1 = rep_len(y, n);

  NumericVector out(n);

  for (int i = 0; i < n; ++i) {
    out[i] = std::min(x1[i], y1[i]);
  }

  return out;
}
```

`f1()` is the same as `mean()`:

```{r}
x <- c(1, 2, 3, 4, 5, 6)

f1(x)
mean(x)
```

`f2()` is the same as `cumsum()`:

```{r}
x <- c(1, 3, 5, 6)

f2(x)
cumsum(x)
```

`f3()` is the same as `any()`:

```{r}
x1 <- c(TRUE, FALSE, FALSE, TRUE)
x2 <- c(FALSE, FALSE)

f3(x1)
any(x1)

f3(x2)
any(x2)
```

`f4()` is the same as `Position()`:

```{r}
x <- list("a", TRUE, "m", 2)

f4(is.numeric, x)
Position(is.numeric, x)
```

`f5()` is the same as `pmin()`:

```{r}
v1 <- c(1, 3, 4, 5, 6, 7)
v2 <- c(1, 2, 7, 2, 8, 1)

f5(v1, v2)
pmin(v1, v2)
```

### Q2. Converting base function to Rcpp {-}

The performance benefits are not going to be observed if the function is primitive since those are already tuned to the max in R for performance. So, expect performance gain only for `diff()` and `var()`.

```{r}
is.primitive(all)
is.primitive(cumprod)
is.primitive(diff)
is.primitive(range)
is.primitive(var)
```

- `all()`

```{Rcpp}
#include <vector>
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
bool allC(std::vector<bool> x)
{
    for (auto xElement : x)
    {
        if (!xElement) return false;
    }

    return true;
}
```

```{r}
v1 <- rep(TRUE, 10)
v2 <- c(rep(TRUE, 5), rep(FALSE, 5))

all(v1)
allC(v1)

all(v2)
allC(v2)

# performance benefits?
bench::mark(
  all(c(rep(TRUE, 1000), rep(FALSE, 1000))),
  allC(c(rep(TRUE, 1000), rep(FALSE, 1000))),
  iterations = 100
)
```

- `cumprod()`

```{Rcpp}
#include <vector>

// [[Rcpp::export]]
std::vector<double> cumulativeProduct(std::vector<double> x)
{
    std::vector<double> out = x;

    for (size_t i = 1; i < x.size(); i++)
    {
        out[i] = out[i - 1] * x[i];
    }

    return out;
}
```


```{r}
v1 <- c(10, 4, 6, 8)

cumprod(v1)
cumulativeProduct(v1)

# performance benefits?
bench::mark(
  cumprod(v1),
  cumulativeProduct(v1),
  iterations = 100
)
```

- `diff()`

- `var()`

```{Rcpp}
#include <vector>
#include <cmath>
#include <numeric>
using namespace std;

// [[Rcpp::export]]
double variance(std::vector<double> x)
{
    double sumSquared{0};

    double mean = std::accumulate(x.begin(), x.end(), 0.0) / x.size();

    for (auto xElement : x)
    {
        sumSquared += pow(xElement - mean, 2.0);
    }

    return sumSquared / (x.size() - 1);
}
```


```{r}
v1 <- c(1, 4, 7, 8)

var(v1)
variance(v1)

# performance benefits?
bench::mark(
  var(v1),
  variance(v1),
  iterations = 100
)
```
