# Rewriting R code in C++

```{r Rcpp-1, include = FALSE}
source("common.R")
```

```{r Rcpp-2, warning=FALSE,message=FALSE}
library(Rcpp, warn.conflicts = FALSE)
```

## Getting started with C++ (Exercises 25.2.6)

**Q1.** With the basics of C++ in hand, it's now a great time to practice by reading and writing some simple C++ functions. For each of the following functions, read the code and figure out what the corresponding base R function is. You might not understand every part of the code yet, but you should be able to figure out the basics of what the function does.

```{Rcpp Rcpp-3}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double f1(NumericVector x) {
  int n = x.size();
  double y = 0;

  for(int i = 0; i < n; ++i) {
    y += x[i] / n;
  }
  return y;
}

// [[Rcpp::export]]
NumericVector f2(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] + x[i];
  }
  return out;
}

// [[Rcpp::export]]
bool f3(LogicalVector x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    if (x[i]) return true;
  }
  return false;
}

// [[Rcpp::export]]
int f4(Function pred, List x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    LogicalVector res = pred(x[i]);
    if (res[0]) return i + 1;
  }
  return 0;
}

// [[Rcpp::export]]
NumericVector f5(NumericVector x, NumericVector y) {
  int n = std::max(x.size(), y.size());
  NumericVector x1 = rep_len(x, n);
  NumericVector y1 = rep_len(y, n);

  NumericVector out(n);

  for (int i = 0; i < n; ++i) {
    out[i] = std::min(x1[i], y1[i]);
  }

  return out;
}
```

**A1.** 

`f1()` is the same as `mean()`:

```{r Rcpp-4}
x <- c(1, 2, 3, 4, 5, 6)

f1(x)
mean(x)
```

`f2()` is the same as `cumsum()`:

```{r Rcpp-5}
x <- c(1, 3, 5, 6)

f2(x)
cumsum(x)
```

`f3()` is the same as `any()`:

```{r Rcpp-6}
x1 <- c(TRUE, FALSE, FALSE, TRUE)
x2 <- c(FALSE, FALSE)

f3(x1)
any(x1)

f3(x2)
any(x2)
```

`f4()` is the same as `Position()`:

```{r Rcpp-7}
x <- list("a", TRUE, "m", 2)

f4(is.numeric, x)
Position(is.numeric, x)
```

`f5()` is the same as `pmin()`:

```{r Rcpp-8}
v1 <- c(1, 3, 4, 5, 6, 7)
v2 <- c(1, 2, 7, 2, 8, 1)

f5(v1, v2)
pmin(v1, v2)
```

**Q2.** To practice your function writing skills, convert the following functions into C++. For now, assume the inputs have no missing values.

1. `all()`.

2. `cumprod()`, `cummin()`, `cummax()`.

3. `diff()`. Start by assuming lag 1, and then generalise for lag `n`.

4. `range()`.

5. `var()`. Read about the approaches you can take on  [Wikipedia](http://en.wikipedia.org/wiki/Algorithms_for_calculating_variance). Whenever implementing a numerical algorithm, it's always good to check what is already known about the problem.

**A2.** The performance benefits are not going to be observed if the function is primitive since those are already tuned to the max in R for performance. So, expect performance gain only for `diff()` and `var()`.

```{r Rcpp-9}
is.primitive(all)
is.primitive(cumprod)
is.primitive(diff)
is.primitive(range)
is.primitive(var)
```

- `all()`

```{Rcpp Rcpp-10}
#include <vector>
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
bool allC(std::vector<bool> x)
{
    for (const auto& xElement : x)
    {
        if (!xElement) return false;
    }

    return true;
}
```

```{r Rcpp-11}
v1 <- rep(TRUE, 10)
v2 <- c(rep(TRUE, 5), rep(FALSE, 5))

all(v1)
allC(v1)

all(v2)
allC(v2)

# performance benefits?
bench::mark(
  all(c(rep(TRUE, 1000), rep(FALSE, 1000))),
  allC(c(rep(TRUE, 1000), rep(FALSE, 1000))),
  iterations = 100
)
```

- `cumprod()`

```{Rcpp Rcpp-12}
#include <vector>

// [[Rcpp::export]]
std::vector<double> cumprodC(const std::vector<double> &x)
{
    std::vector<double> out{x};

    for (std::size_t i = 1; i < x.size(); i++)
    {
        out[i] = out[i - 1] * x[i];
    }

    return out;
}
```


```{r Rcpp-13}
v1 <- c(10, 4, 6, 8)

cumprod(v1)
cumprodC(v1)

# performance benefits?
bench::mark(
  cumprod(v1),
  cumprodC(v1),
  iterations = 100
)
```

- `cumminC()`

```{Rcpp Rcpp-14}
#include <vector>
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> cumminC(const std::vector<double> &x)
{
    std::vector<double> out{x};

    for (std::size_t i = 1; i < x.size(); i++)
    {
        out[i] = (out[i] < out[i - 1]) ? out[i] : out[i - 1];
    }

    return out;
}
```

```{r Rcpp-15}
v1 <- c(3:1, 2:0, 4:2)

cummin(v1)
cumminC(v1)

# performance benefits?
bench::mark(
  cummin(v1),
  cumminC(v1),
  iterations = 100
)
```

- `cummaxC()`

```{Rcpp Rcpp-16}
#include <vector>
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> cummaxC(const std::vector<double> &x)
{
    std::vector<double> out{x};

    for (std::size_t i = 1; i < x.size(); i++)
    {
        out[i] = (out[i] > out[i - 1]) ? out[i] : out[i - 1];
    }
    
    return out;
}
```

```{r Rcpp-17}
v1 <- c(3:1, 2:0, 4:2)

cummax(v1)
cummaxC(v1)

# performance benefits?
bench::mark(
  cummax(v1),
  cummaxC(v1),
  iterations = 100
)
```

- `diff()`

```{Rcpp Rcpp-18}
#include <vector>
#include <functional>
#include <algorithm>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> diffC(const std::vector<double> &x, int lag)
{
    std::vector<double> vec_start;
    std::vector<double> vec_lagged;
    std::vector<double> vec_diff;

    for (std::size_t i = lag; i < x.size(); i++)
    {
        vec_lagged.push_back(x[i]);
    }

    for (std::size_t i = 0; i < (x.size() - lag); i++)
    {
        vec_start.push_back(x[i]);
    }

    std::transform(
        vec_lagged.begin(), vec_lagged.end(),
        vec_start.begin(), std::back_inserter(vec_diff),
        std::minus<double>());

    return vec_diff;
}
```

```{r Rcpp-19}
v1 <- c(1, 2, 4, 8, 13)
v2 <- c(1, 2, NA, 8, 13)

diff(v1, 2)
diffC(v1, 2)

diff(v2, 2)
diffC(v2, 2)

# performance benefits?
bench::mark(
  diff(v1, 2),
  diffC(v1, 2),
  iterations = 100
)
```

- `range()`

```{Rcpp Rcpp-20}
#include <iostream>
#include <vector>
#include <algorithm>
using namespace std;

// [[Rcpp::export]]
std::vector<double> rangeC(std::vector<double> x)
{
    std::vector<double> rangeVec{0.0, 0.0};

    rangeVec.at(0) = *std::min_element(x.begin(), x.end());
    rangeVec.at(1) = *std::max_element(x.begin(), x.end());

    return rangeVec;
}
```

```{r Rcpp-21}
v1 <- c(10, 4, 6, 8)

range(v1)
rangeC(v1)

# performance benefits?
bench::mark(
  range(v1),
  rangeC(v1),
  iterations = 100
)
```

- `var()`

```{Rcpp Rcpp-22}
#include <vector>
#include <cmath>
#include <numeric>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
double variance(std::vector<double> x)
{
    double sumSquared{0};

    double mean = std::accumulate(x.begin(), x.end(), 0.0) / x.size();

    for (const auto& xElement : x)
    {
        sumSquared += pow(xElement - mean, 2.0);
    }

    return sumSquared / (x.size() - 1);
}
```


```{r Rcpp-23}
v1 <- c(1, 4, 7, 8)

var(v1)
variance(v1)

# performance benefits?
bench::mark(
  var(v1),
  variance(v1),
  iterations = 100
)
```

## Missing values (Exercises 25.4.5)

**Q1.** Rewrite any of the functions from Exercise 25.2.6 to deal with missing values. If `na.rm` is true, ignore the missing values. If `na.rm` is false,  return a missing value if the input contains any missing values. Some good functions to practice with are `min()`, `max()`, `range()`, `mean()`, and `var()`.

**A1.** We will only create a version of `range()` that deals with missing values. The same principle applies to others:

```{Rcpp Rcpp-24}
#include <iostream>
#include <vector>
#include <algorithm>
#include <math.h>
#include <Rcpp.h>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> rangeC_NA(std::vector<double> x, bool removeNA = true)
{
    std::vector<double> rangeVec{0.0, 0.0};

    bool naPresent = std::any_of(
        x.begin(),
        x.end(),
        [](double d)
        { return isnan(d); });

    if (naPresent)
    {
        if (removeNA)
        {
            std::remove(x.begin(), x.end(), NAN);
        }
        else
        {
            rangeVec.at(0) = NA_REAL; // NAN;
            rangeVec.at(1) = NA_REAL; // NAN;

            return rangeVec;
        }
    }

    rangeVec.at(0) = *std::min_element(x.begin(), x.end());
    rangeVec.at(1) = *std::max_element(x.begin(), x.end());

    return rangeVec;
}
```

```{r Rcpp-25}
v1 <- c(10, 4, NA, 6, 8)

range(v1, na.rm = FALSE)
rangeC_NA(v1, FALSE)

range(v1, na.rm = TRUE)
rangeC_NA(v1, TRUE)
```

**Q2.** Rewrite `cumsum()` and `diff()` so they can handle missing values. Note that these functions have slightly more complicated behaviour.

**A2.** The `cumsum()` docs say:

> An `NA` value in `x` causes the corresponding and following elements of the return value to be `NA`, as does integer overflow in cumsum (with a warning).

Similarly, `diff()` docs say:

> `NA`'s propagate.

Therefore, both of these functions don't allow removing missing values and the `NA`s propagate.

As seen from the examples above, `diffC()` already behaves this way.

Similarly, `cumsumC()` propagates `NA`s as well.

```{Rcpp Rcpp-26}
#include <Rcpp.h>
using namespace Rcpp;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
NumericVector cumsumC(NumericVector x) {
  int n = x.size();
  NumericVector out(n);
  
  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] + x[i];
  }
  
  return out;
}
```

```{r Rcpp-27}
v1 <- c(1, 2, 3, 4)
v2 <- c(1, 2, NA, 4)

cumsum(v1)
cumsumC(v1)

cumsum(v2)
cumsumC(v2)
```

## Standard Template Library (Exercises 25.5.7)

**Q1.** To practice using the STL algorithms and data structures, implement the following using R functions in C++, using the hints provided:

**A1.** 

1. `median.default()` using `partial_sort`.

```{Rcpp Rcpp-28}
#include <iostream>
#include <vector>
#include <algorithm>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
double medianC(std::vector<double> &x)
{
    int middleIndex = static_cast<int>(x.size() / 2);

    std::partial_sort(x.begin(), x.begin() + middleIndex, x.end());

    // for even number of observations
    if (x.size() % 2 == 0)
    {
        return (x[middleIndex - 1] + x[middleIndex]) / 2;
    }

    return x[middleIndex];
}
```


```{r Rcpp-29}
v1 <- c(1, 3, 3, 6, 7, 8, 9)
v2 <- c(1, 2, 3, 4, 5, 6, 8, 9)

median.default(v1)
medianC(v1)

median.default(v2)
medianC(v2)

# performance benefits?
bench::mark(
  median.default(v2),
  medianC(v2),
  iterations = 100
)
```

1. `%in%` using `unordered_set` and the `find()` or `count()` methods.

```{Rcpp Rcpp-30}
#include <vector>
#include <unordered_set>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<bool> matchC(const std::vector<double> &x, const std::vector<double> &table)
{
    std::unordered_set<double> tableUnique(table.begin(), table.end());
    std::vector<bool> out;

    for (const auto &xElem : x)
    {
        out.push_back(tableUnique.find(xElem) != tableUnique.end() ? true : false);
    }

    return out;
}
```

```{r Rcpp-31}
x1 <- c(3, 4, 8)
x2 <- c(1, 2, 3, 3, 4, 4, 5, 6)

x1 %in% x2
matchC(x1, x2)

# performance benefits?
bench::mark(
  x1 %in% x2,
  matchC(x1, x2),
  iterations = 100
)
```

1. `unique()` using an `unordered_set` (challenge: do it in one line!).

```{Rcpp Rcpp-32}
#include <unordered_set>
#include <vector>
#include <iostream>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::unordered_set<double> uniqueC(const std::vector<double> &x)
{
    std::unordered_set<double> xSet(x.begin(), x.end());

    return xSet;
}
```

Note that these functions are **not** comparable. As far as I can see, there is no way to get the same output as the R version of the function using the `unordered_set` data structure.

```{r Rcpp-33}
v1 <- c(1, 3, 3, 6, 7, 8, 9)

unique(v1)
uniqueC(v1)
```

We can make comparable version using `set` data structure:

```{Rcpp Rcpp-34}
#include <set>
#include <vector>
#include <iostream>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::set<double> uniqueC2(const std::vector<double> &x)
{
    std::set<double> xSet(x.begin(), x.end());

    return xSet;
}
```

```{r Rcpp-35}
v1 <- c(1, 3, 3, 6, 7, 8, 9)

unique(v1)
uniqueC2(v1)

# performance benefits?
bench::mark(
  unique(v1),
  uniqueC2(v1),
  iterations = 100
)
```

1. `min()` using `std::min()`, or `max()` using `std::max()`.

```{Rcpp Rcpp-36}
#include <iostream>
#include <vector>
#include <algorithm>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
const double minC(const std::vector<double> &x)
{
     return *std::min_element(x.begin(), x.end());
}

// [[Rcpp::export]]
const double maxC(std::vector<double> x)
{
     return *std::max_element(x.begin(), x.end());
}
```

```{r Rcpp-37}
v1 <- c(3, 3, 6, 1, 9, 7, 8)

min(v1)
minC(v1)

# performance benefits?
bench::mark(
  min(v1),
  minC(v1),
  iterations = 100
)

max(v1)
maxC(v1)

# performance benefits?
bench::mark(
  max(v1),
  maxC(v1),
  iterations = 100
)
```

1. `which.min()` using `min_element`, or `which.max()` using `max_element`.

```{Rcpp Rcpp-38}
#include <vector>
#include <algorithm>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
int which_maxC(std::vector<double> &x)
{
    int maxIndex = std::distance(x.begin(), std::max_element(x.begin(), x.end()));
  
    // R is 1-index based, while C++ is 0-index based
    return maxIndex + 1;
}

// [[Rcpp::export]]
int which_minC(std::vector<double> &x)
{
    int minIndex = std::distance(x.begin(), std::min_element(x.begin(), x.end()));
  
    // R is 1-index based, while C++ is 0-index based
    return minIndex + 1;
}
```


```{r Rcpp-39}
v1 <- c(3, 3, 6, 1, 9, 7, 8)

which.min(v1)
which_minC(v1)

# performance benefits?
bench::mark(
  which.min(v1),
  which_minC(v1),
  iterations = 100
)

which.max(v1)
which_maxC(v1)

# performance benefits?
bench::mark(
  which.max(v1),
  which_maxC(v1),
  iterations = 100
)
```

1. `setdiff()`, `union()`, and `intersect()` for integers using sorted ranges and `set_union`, `set_intersection` and `set_difference`.

Note that the following C++ implementations of given functions are not strictly equivalent to their R versions. As far as I can see, there is no way for them to be identical while satisfying the specifications mentioned in the question.

- `union()`

```{Rcpp Rcpp-40}
#include <algorithm>
#include <iostream>
#include <vector>
#include <set>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::set<int> unionC(std::vector<int> &v1, std::vector<int> &v2)
{
    std::sort(v1.begin(), v1.end());
    std::sort(v2.begin(), v2.end());

    std::vector<int> union_vec(v1.size() + v2.size());
    auto it = std::set_union(v1.begin(), v1.end(), v2.begin(), v2.end(), union_vec.begin());

    union_vec.resize(it - union_vec.begin());
    std::set<int> union_set(union_vec.begin(), union_vec.end());

    return union_set;
}
```

```{r Rcpp-41}
v1 <- c(1, 4, 5, 5, 5, 6, 2)
v2 <- c(4, 1, 6, 8)

union(v1, v2)
unionC(v1, v2)
```

- `intersect()`

```{Rcpp Rcpp-42}
#include <algorithm>
#include <iostream>
#include <vector>
#include <set>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::set<int> intersectC(std::vector<int> &v1, std::vector<int> &v2)
{
    std::sort(v1.begin(), v1.end());
    std::sort(v2.begin(), v2.end());

    std::vector<int> union_vec(v1.size() + v2.size());
    auto it = std::set_intersection(v1.begin(), v1.end(), v2.begin(), v2.end(), union_vec.begin());

    union_vec.resize(it - union_vec.begin());
    std::set<int> union_set(union_vec.begin(), union_vec.end());

    return union_set;
}
```

```{r Rcpp-43}
v1 <- c(1, 4, 5, 5, 5, 6, 2)
v2 <- c(4, 1, 6, 8)

intersect(v1, v2)
intersectC(v1, v2)
```

- `setdiff()`

```{Rcpp Rcpp-44}
#include <algorithm>
#include <iostream>
#include <vector>
#include <set>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::set<int> setdiffC(std::vector<int> &v1, std::vector<int> &v2)
{
    std::sort(v1.begin(), v1.end());
    std::sort(v2.begin(), v2.end());

    std::vector<int> union_vec(v1.size() + v2.size());
    auto it = std::set_difference(v1.begin(), v1.end(), v2.begin(), v2.end(), union_vec.begin());

    union_vec.resize(it - union_vec.begin());
    std::set<int> union_set(union_vec.begin(), union_vec.end());

    return union_set;
}
```

```{r Rcpp-45}
v1 <- c(1, 4, 5, 5, 5, 6, 2)
v2 <- c(4, 1, 6, 8)

setdiff(v1, v2)
setdiffC(v1, v2)
```

## Session information

```{r Rcpp-46}
sessioninfo::session_info(include_base = TRUE)
```

