# Rewriting R code in C++

```{r, include = FALSE}
source("common.R")
```

## Exercises 25.2.6

**Q1.** With the basics of C++ in hand, it's now a great time to practice by reading and writing some simple C++ functions. For each ofthe following functions, read the code and figure out what the corresponding base R function is. You might not understand every part of the code yet, but you should be able to figure out the basics of what the function does.
 
```{r}
library(Rcpp)
```

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double f1(NumericVector x) {
  int n = x.size();
  double y = 0;

  for(int i = 0; i < n; ++i) {
    y += x[i] / n;
  }
  return y;
}

// [[Rcpp::export]]
NumericVector f2(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] + x[i];
  }
  return out;
}

// [[Rcpp::export]]
bool f3(LogicalVector x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    if (x[i]) return true;
  }
  return false;
}

// [[Rcpp::export]]
int f4(Function pred, List x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    LogicalVector res = pred(x[i]);
    if (res[0]) return i + 1;
  }
  return 0;
}

// [[Rcpp::export]]
NumericVector f5(NumericVector x, NumericVector y) {
  int n = std::max(x.size(), y.size());
  NumericVector x1 = rep_len(x, n);
  NumericVector y1 = rep_len(y, n);

  NumericVector out(n);

  for (int i = 0; i < n; ++i) {
    out[i] = std::min(x1[i], y1[i]);
  }

  return out;
}
```

**A1.** 

`f1()` is the same as `mean()`:

```{r}
x <- c(1, 2, 3, 4, 5, 6)

f1(x)
mean(x)
```

`f2()` is the same as `cumsum()`:

```{r}
x <- c(1, 3, 5, 6)

f2(x)
cumsum(x)
```

`f3()` is the same as `any()`:

```{r}
x1 <- c(TRUE, FALSE, FALSE, TRUE)
x2 <- c(FALSE, FALSE)

f3(x1)
any(x1)

f3(x2)
any(x2)
```

`f4()` is the same as `Position()`:

```{r}
x <- list("a", TRUE, "m", 2)

f4(is.numeric, x)
Position(is.numeric, x)
```

`f5()` is the same as `pmin()`:

```{r}
v1 <- c(1, 3, 4, 5, 6, 7)
v2 <- c(1, 2, 7, 2, 8, 1)

f5(v1, v2)
pmin(v1, v2)
```

**Q2.** To practice your function writing skills, convert the following functions into C++. For now, assume the inputs have no missing values.

1. `all()`.

2. `cumprod()`, `cummin()`, `cummax()`.

3. `diff()`. Start by assuming lag 1, and then generalise for lag `n`.

4. `range()`.

5. `var()`. Read about the approaches you can take on  [Wikipedia](http://en.wikipedia.org/wiki/Algorithms_for_calculating_variance). Whenever implementing a numerical algorithm, it's always good to check what is already known about the problem.

**A2.** The performance benefits are not going to be observed if the function is primitive since those are already tuned to the max in R for performance. So, expect performance gain only for `diff()` and `var()`.

```{r}
is.primitive(all)
is.primitive(cumprod)
is.primitive(diff)
is.primitive(range)
is.primitive(var)
```

- `all()`

```{Rcpp}
#include <vector>
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
bool allC(std::vector<bool> x)
{
    for (const auto& xElement : x)
    {
        if (!xElement) return false;
    }

    return true;
}
```

```{r}
v1 <- rep(TRUE, 10)
v2 <- c(rep(TRUE, 5), rep(FALSE, 5))

all(v1)
allC(v1)

all(v2)
allC(v2)

# performance benefits?
bench::mark(
  all(c(rep(TRUE, 1000), rep(FALSE, 1000))),
  allC(c(rep(TRUE, 1000), rep(FALSE, 1000))),
  iterations = 100
)
```

- `cumprod()`

```{Rcpp}
#include <vector>

// [[Rcpp::export]]
std::vector<double> cumprodC(const std::vector<double> &x)
{
    std::vector<double> out{x};

    for (size_t i = 1; i < x.size(); i++)
    {
        out[i] = out[i - 1] * x[i];
    }

    return out;
}
```


```{r}
v1 <- c(10, 4, 6, 8)

cumprod(v1)
cumprodC(v1)

# performance benefits?
bench::mark(
  cumprod(v1),
  cumprodC(v1),
  iterations = 100
)
```

- `cumminC()`

```{Rcpp}
#include <vector>
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> cumminC(const std::vector<double> &x)
{
    std::vector<double> out{x};

    for (size_t i = 1; i < x.size(); i++)
    {
        out[i] = (out[i] < out[i - 1]) ? out[i] : out[i - 1];
    }

    return out;
}
```

```{r}
v1 <- c(3:1, 2:0, 4:2)

cummin(v1)
cumminC(v1)

# performance benefits?
bench::mark(
  cummin(v1),
  cumminC(v1),
  iterations = 100
)
```

- `cummaxC()`

```{Rcpp}
#include <vector>
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> cummaxC(const std::vector<double> &x)
{
    std::vector<double> out{x};

    for (size_t i = 1; i < x.size(); i++)
    {
        out[i] = (out[i] > out[i - 1]) ? out[i] : out[i - 1];
    }
    
    return out;
}
```

```{r}
v1 <- c(3:1, 2:0, 4:2)

cummax(v1)
cummaxC(v1)

# performance benefits?
bench::mark(
  cummax(v1),
  cummaxC(v1),
  iterations = 100
)
```

- `diff()`

```{Rcpp}
#include <vector>
#include <functional>
#include <algorithm>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> diffC(const std::vector<double> &x, int lag)
{
    std::vector<double> vec_start;
    std::vector<double> vec_lagged;

    for (size_t i = lag; i < x.size(); i++)
    {
        vec_lagged.push_back(x[i]);
    }

    for (size_t i = 0; i < (x.size() - lag); i++)
    {
        vec_start.push_back(x[i]);
    }

    std::transform(
        vec_lagged.begin(), vec_lagged.end(),
        vec_start.begin(), vec_lagged.begin(),
        std::minus<double>());

    return vec_lagged;
}
```

```{r}
v1 <- c(1, 2, 4, 8, 13)

diff(v1, 2)
diffC(v1, 2)

# performance benefits?
bench::mark(
  diff(v1, 2),
  diffC(v1, 2),
  iterations = 100
)
```

- `range()`

```{Rcpp}
#include <iostream>
#include <vector>
#include <algorithm>
using namespace std;

// [[Rcpp::export]]
std::vector<double> rangeC(std::vector<double> x)
{
    std::vector<double> rangeVec{0.0, 0.0};

    rangeVec.at(0) = *std::min_element(x.begin(), x.end());
    rangeVec.at(1) = *std::max_element(x.begin(), x.end());

    return rangeVec;
}
```

```{r}
v1 <- c(10, 4, 6, 8)

range(v1)
rangeC(v1)

# performance benefits?
bench::mark(
  range(v1),
  rangeC(v1),
  iterations = 100
)
```

- `var()`

```{Rcpp}
#include <vector>
#include <cmath>
#include <numeric>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
double variance(std::vector<double> x)
{
    double sumSquared{0};

    double mean = std::accumulate(x.begin(), x.end(), 0.0) / x.size();

    for (const auto& xElement : x)
    {
        sumSquared += pow(xElement - mean, 2.0);
    }

    return sumSquared / (x.size() - 1);
}
```


```{r}
v1 <- c(1, 4, 7, 8)

var(v1)
variance(v1)

# performance benefits?
bench::mark(
  var(v1),
  variance(v1),
  iterations = 100
)
```

## Exercises 25.4.5

**Q1.** Rewrite any of the functions from Exercise 25.2.6 to deal with missing values. If `na.rm` is true, ignore the missing values. If `na.rm` is false,  return a missing value if the input contains any missing values. Some good functions to practice with are `min()`, `max()`, `range()`, `mean()`, and `var()`.

**A1.**

```{Rcpp}
#include <iostream>
#include <vector>
#include <algorithm>
#include <math.h>
#include <Rcpp.h>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::vector<double> rangeC_NA(std::vector<double> x, bool removeNA = true)
{
    std::vector<double> rangeVec{0.0, 0.0};

    bool naPresent = std::any_of(
        x.begin(),
        x.end(),
        [](double d)
        { return isnan(d); });

    if (naPresent)
    {
        if (removeNA)
        {
            std::remove(x.begin(), x.end(), NAN);
        }
        else
        {
            rangeVec.at(0) = NA_REAL; // NAN;
            rangeVec.at(1) = NA_REAL; // NAN;

            return rangeVec;
        }
    }

    rangeVec.at(0) = *std::min_element(x.begin(), x.end());
    rangeVec.at(1) = *std::max_element(x.begin(), x.end());

    return rangeVec;
}
```

```{r}
v1 <- c(10, 4, NA, 6, 8)

range(v1, na.rm = FALSE)
rangeC_NA(v1, FALSE)

range(v1, na.rm = TRUE)
rangeC_NA(v1, TRUE)
```

**Q2.** Rewrite `cumsum()` and `diff()` so they can handle missing values. Note that these functions have slightly more complicated behaviour.

## Exercises 25.5.7

**Q1.** To practice using the STL algorithms and data structures, implement the following using R functions in C++, using the hints provided:

**A1.** 

1. `median.default()` using `partial_sort`.

```{Rcpp}
#include <iostream>
#include <vector>
#include <algorithm>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
double medianC(std::vector<double> &x)
{
    int middleIndex = static_cast<int>(x.size() / 2);

    std::partial_sort(x.begin(), x.begin() + middleIndex, x.end());

    // for even number of observations
    if (x.size() % 2 == 0)
    {
        return (x[middleIndex - 1] + x[middleIndex]) / 2;
    }

    return x[middleIndex];
}
```


```{r}
v1 <- c(1, 3, 3, 6, 7, 8, 9)
v2 <- c(1, 2, 3, 4, 5, 6, 8, 9)

median.default(v1)
medianC(v1)

median.default(v2)
medianC(v2)

# performance benefits?
bench::mark(
  median.default(v2),
  medianC(v2),
  iterations = 100
)
```

1. `%in%` using `unordered_set` and the `find()` or `count()` methods.

1. `unique()` using an `unordered_set` (challenge: do it in one line!).

```{Rcpp}
#include <unordered_set>
#include <vector>
#include <iostream>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
std::unordered_set<double> uniqueC(const std::vector<double> &x)
{
    std::unordered_set<double> xSet(x.begin(), x.end());

    return xSet;
}
```

```{r}
v1 <- c(1, 3, 3, 6, 7, 8, 9)

unique(v1)
uniqueC(v1)

# performance benefits?
# bench::mark(
#   unique(v1),
#   uniqueC(v1),
#   iterations = 100
# )
```

1. `min()` using `std::min()`, or `max()` using `std::max()`.

```{Rcpp}
#include <iostream>
#include <vector>
#include <algorithm>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
const double minC(const std::vector<double> &x)
{
     return *std::min_element(x.begin(), x.end());
}

// [[Rcpp::export]]
const double maxC(std::vector<double> x)
{
     return *std::max_element(x.begin(), x.end());
}
```

```{r}
v1 <- c(3, 3, 6, 1, 9, 7, 8)

min(v1)
minC(v1)

# performance benefits?
bench::mark(
  min(v1),
  minC(v1),
  iterations = 100
)

max(v1)
maxC(v1)

# performance benefits?
bench::mark(
  max(v1),
  maxC(v1),
  iterations = 100
)
```

1. `which.min()` using `min_element`, or `which.max()` using `max_element`.

```{Rcpp}
#include <vector>
#include <algorithm>
#include <iostream>
using namespace std;
// [[Rcpp::plugins(cpp11)]]

// [[Rcpp::export]]
int which_maxC(std::vector<double> &x)
{
    int maxIndex = std::distance(x.begin(), std::max_element(x.begin(), x.end()));
  
    // R is 1-index based, while C++ is 0-index based
    return maxIndex + 1;
}

// [[Rcpp::export]]
int which_minC(std::vector<double> &x)
{
    int minIndex = std::distance(x.begin(), std::min_element(x.begin(), x.end()));
  
    // R is 1-index based, while C++ is 0-index based
    return minIndex + 1;
}
```


```{r}
v1 <- c(3, 3, 6, 1, 9, 7, 8)

which.min(v1)
which_minC(v1)

# performance benefits?
bench::mark(
  which.min(v1),
  which_minC(v1),
  iterations = 100
)

which.max(v1)
which_maxC(v1)

# performance benefits?
bench::mark(
  which.max(v1),
  which_maxC(v1),
  iterations = 100
)
```

1. `setdiff()`, `union()`, and `intersect()` for integers using sorted ranges and `set_union`, `set_intersection` and `set_difference`.

