# S3

## Exercise 13.2.1

### Q1. Differences between `t.test` and `t.data.frame`

```{r}
library(sloop)

# function type
ftype(t.test)
ftype(t.data.frame)
```

- `t.test()` is a **generic** function to perform t-test.

- `t.data.frame` is a **method** for generic `t()` (a matrix transform function) and will be dispatched for `data.frame` objects that need to be transformed.

### Q2. base-R function with `.`

- `all.equal()`
- Most of `as.*` functions like `as.data.frame()`
- `install.packages()`
etc.

For example,

```{r}
ftype(as.data.frame)
```

### Q3. What does `as.data.frame.data.frame()` do?

It's a **method** for generic `as.data.frame()`.

Less confusing: `asDataFrame.DataFrame()`.

### Q4. Difference in behavior

Before unclassing, the S3 dispatches `.Date` method, while after `.numeric` method.

Before

```{r}
some_days <- as.Date("2017-01-31") + sample(10, 5)

some_days

s3_dispatch(mean(some_days))

mean(some_days)
```

After

```{r}
unclass(some_days)

mean(unclass(some_days))

s3_dispatch(mean(unclass(some_days)))
```

### Q5. Object properties 

```{r}
x <- ecdf(rpois(100, 10))
x

otype(x)

attributes(x)

s3_class(x)
```

### Q6. Object properties 

```{r}
x <- table(rpois(100, 5))
x

otype(x)

attributes(x)

s3_class(x)
```
