# S3

```{r, include = FALSE}
source("common.R")
```

## Exercises 13.2.1 

**Q1.** Describe the difference between `t.test()` and `t.data.frame()`. When is each function called?

**A1.** 

- `t.test()` is a **generic** function to perform a *t*-test.

- `t.data.frame` is a **method** for generic `t()` (a matrix transform function) and will be dispatched for `data.frame` class objects/instances that need to be transformed.

```{r}
library(sloop)

# function type
ftype(t.test)
ftype(t.data.frame)
```

**Q2.** Make a list of commonly used base R functions that contain `.` in their name but are not `S3` methods.

**A2.** Here are a few common R functions with `.` but that are not `S3` methods:

- `all.equal()`
- Most of `as.*` functions (like `as.data.frame()`, `as.numeric()`, etc.)
- `install.packages()`
- `on.exit()`
etc.

For full list, you could do:

```{r, eval=FALSE}
base_functions <- getNamespaceExports("base")

base_functions[grepl("(\\w+)(\\.)(\\w+)", base_functions)]
```

For example,

```{r}
ftype(as.data.frame)
ftype(on.exit)
```

**Q3.** What does the `as.data.frame.data.frame()` method do? Why is it confusing? How could you avoid this confusion in your own code?
 
**A3.** It's an `S3` **method** for **generic** `as.data.frame()`.

```{r}
ftype(as.data.frame.data.frame)
```

It can be seen in all methods supported by this generic:

```{r}
s3_methods_generic("as.data.frame") %>%
  dplyr::filter(class == "data.frame")
```


**Q4.** Describe the difference in behaviour in these two calls.

```{r}
set.seed(1014)
some_days <- as.Date("2017-01-31") + sample(10, 5)
mean(some_days)
mean(unclass(some_days))
```

**A4.** 

- Before unclassing, the `mean` generic dispatches `.Date` method:

```{r}
some_days <- as.Date("2017-01-31") + sample(10, 5)

some_days

s3_dispatch(mean(some_days))

mean(some_days)
```

- After unclassing, the `mean` generic dispatches `.numeric` method:

```{r}
unclass(some_days)

mean(unclass(some_days))

s3_dispatch(mean(unclass(some_days)))
```

**Q5.** What class of object does the following code return? What base type is it built on? What attributes does it use?

```{r, eval=FALSE}
x <- ecdf(rpois(100, 10))
x
```

**A5.** The object is based on base type `closure`^[of "object of type 'closure' is not subsettable" fame], which is a type of function.

```{r}
x <- ecdf(rpois(100, 10))
x

otype(x)
typeof(x)
```

Its class is `ecdf`, which has other superclasses.

```{r}
s3_class(x)
```

Apart from `class`, it has the following attributes:

```{r}
attributes(x)
```

**Q6.** What class of object does the following code return? What base type is it built on? What attributes does it use?

```{r, eval = FALSE}
x <- table(rpois(100, 5))
x
```

**A6.** The object is based on base type `integer`.

```{r}
x <- table(rpois(100, 5))
x

otype(x)
typeof(x)
```

Its class is `table`.

```{r}
s3_class(x)
```

Apart from `class`, it has the following attributes:

```{r}
attributes(x)
```

## Exercises 13.3.4

**Q1.** Write a constructor for `data.frame` objects. What base type is a data frame built on? What attributes does it use? What are the restrictions placed on the individual elements? What about the names?

**A1.** 

```{r}
my_data_frame <- function(...,
                          row.names = NULL,
                          check.rows = FALSE,
                          check.names = TRUE,
                          fix.empty.names = TRUE,
                          stringsAsFactors = FALSE) {
  structure(
    df,
    class = "data.frame"
  )
}
```


**Q2.** Enhance my `factor()` helper to have better behaviour when one or more `values` is not found in `levels`. What does `base::factor()` do in this situation?

**Q3.** Carefully read the source code of `factor()`. What does it do that my constructor does not?
    
**Q4.** Factors have an optional "contrasts" attribute. Read the help for `C()`, and briefly describe the purpose of the attribute. What type should it have? Rewrite the `new_factor()` constructor to include this attribute.

**Q5.** Read the documentation for `utils::as.roman()`. How would you write a constructor for this class? Does it need a validator? What might a helper do?


## Exercises 13.4.4

**Q1.** Read the source code for `t()` and `t.test()` and confirm that `t.test()` is an S3 generic and not an S3 method. What happens if you create an object with class `test` and call `t()` with it? Why?

```{r, results = FALSE}
x <- structure(1:10, class = "test")
t(x)
```

**A1.** Looking at source code of these functions, we can see that both of these are generic, and we can confirm the same using `{sloop}`:

```{r}
t
sloop::is_s3_generic("t")

t.test
sloop::is_s3_generic("t.test")
```

Looking at the `S3` dispatch, we can see that since R can't find `S3` method for `test` class for generic function `t()`, it dispatches the default method, which converts the structure to a matrix:

```{r}
x <- structure(1:10, class = "test")
t(x)
s3_dispatch(t(x))
```

The same behaviour can be observed with a vector:

```{r}
t(1:10)
```

**Q2.** What generics does the `table` class have methods for?

**A2.** The `table` class have methods for the following generics:

```{r}
s3_methods_class("table")
```

**Q3.** What generics does the `ecdf` class have methods for?

**A3.** The `ecdf` class have methods for the following generics:

```{r}
s3_methods_class("ecdf")
```

**Q4.** Which base generic has the greatest number of defined methods?

**A4.** To answer this question, first, let's list all functions base has and only retain the generics.

```{r}
library(purrr)
library(dplyr)

# getting all functions names
objs <- mget(ls("package:base", all = TRUE), inherits = TRUE)
funs <- Filter(is.function, objs)

# extracting only generics
genFuns <- names(funs) %>%
  purrr::keep(~ sloop::is_s3_generic(.x))
```

Now it's a simple matter of counting number of methods per generic and ordering the data frame in descending order of this count:

```{r}
purrr::map_dfr(
  genFuns,
  ~ s3_methods_generic(.)
) %>%
  dplyr::group_by(generic) %>%
  dplyr::tally() %>%
  dplyr::arrange(desc(n))
```

This reveals that the base generic function with most methods is `print()`.

**Q5.** Carefully read the documentation for `UseMethod()` and explain why the following code returns the results that it does. What two usual rules of function evaluation does `UseMethod()` violate?

```{r}
g <- function(x) {
  x <- 10
  y <- 10
  UseMethod("g")
}
g.default <- function(x) c(x = x, y = y)
x <- 1
y <- 1
g(x)
```

**Q6.** What are the arguments to `[`? Why is this a hard question to answer?

**A6.** It is difficult to say how many formal arguments the subsetting `[` operator has because it is a generic function with methods for vectors, matrices, arrays, lists, etc., and these different methods have different number of arguments:

```{r}
s3_methods_generic("[") %>%
  dplyr::filter(source == "base")
```

We can sample a few of them to see the diversity in the number of formal arguments:

```{r}
# table
names(formals(`[.table`))

# Date
names(formals(`[.Date`))

# data frame
names(formals(`[.data.frame`))

# etc.
```

## Exercises 13.5.1

**Q1.** Categorise the objects returned by `lm()`, `factor()`, `table()`, `as.Date()`, `as.POSIXct()` `ecdf()`, `ordered()`, `I()` into the styles described above.

**Q2.** What would a constructor function for `lm` objects, `new_lm()`, look like? Use `?lm` and experimentation to figure out the required fields and their types.

## Exercises 13.6.3

**Q1.** How does `[.Date` support subclasses? How does it fail to support subclasses?

**Q2.** R has two classes for representing date time data, `POSIXct` and `POSIXlt`, which both inherit from `POSIXt`. Which generics have different behaviours for the two classes? Which generics share the same behaviour?

**Q3.** What do you expect this code to return? What does it actually return? Why?

```{r, eval = FALSE}
generic2 <- function(x) UseMethod("generic2")
generic2.a1 <- function(x) "a1"
generic2.a2 <- function(x) "a2"
generic2.b <- function(x) {
  class(x) <- "a1"
  NextMethod()
}
generic2(structure(list(), class = c("b", "a2")))
```

## Exercises 13.7.5

**Q1.** Explain the differences in dispatch below:

```{r}
length.integer <- function(x) 10

x1 <- 1:5
class(x1)
s3_dispatch(length(x1))

x2 <- structure(x1, class = "integer")
class(x2)
s3_dispatch(length(x2))
```

**Q2.** What classes have a method for the `Math` group generic in base R? Read the source code. How do the methods work?

**Q3.** `Math.difftime()` is more complicated than I described. Why?
