# Subsetting

## Exercise 4.2.6

**Q1.** Fix each of the following common data frame subsetting errors:

```{r, eval = FALSE}
mtcars[mtcars$cyl = 4, ]
mtcars[-1:4, ]
mtcars[mtcars$cyl <= 5]
mtcars[mtcars$cyl == 4 | 6, ]
```

**A1.** Fixed versions of these commands:

```{r, eval=FALSE}
mtcars[mtcars$cyl == 4, ]
mtcars[-(1:4), ]
mtcars[mtcars$cyl <= 5, ]
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
```

**Q2.** Why does the following code yield five missing values?

```{r}
x <- 1:5
x[NA]
```

**A2.** This is because of two reasons:

- The default type of `NA` in R is of `logical` type.

```{r}
typeof(NA)
```

- R recycles indexes to match the length of the vector.

```{r}
x <- 1:5
x[c(TRUE, FALSE)] # recycled to c(TRUE, FALSE, TRUE, FALSE, TRUE)
```

**Q3.** What does `upper.tri()` return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?

```{r, eval = FALSE}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```

A3. The documentation for `upper.tri()` states-

> Returns a matrix of logicals the same size of a given matrix with entries `TRUE` in the **upper triangle**

That is, `upper.tri()` return a matrix of logicals.

```{r}
(x <- outer(1:5, 1:5, FUN = "*"))

upper.tri(x)
```

When used with a matrix for subsetting, this logical matrix returns a vector:

```{r}
x[upper.tri(x)]
```

**Q4.**  Why does `mtcars[1:20]` return an error? How does it differ from the similar `mtcars[1:20, ]`?

When indexed like a list, data frame columns at given indices will be selected.

```{r}
head(mtcars[1:2])
```

`mtcars[1:20]` doesn't work because there are `r length(mtcars)` columns in `mtcars` dataset.

On the other hand, `mtcars[1:20, ]` indexes a dataframe like a matrix, and because there are indeed 20 rows in `mtcars`, all columns with these rows are selected.

```{r}
nrow(mtcars[1:20, ])
```

**Q5.** Implement your own function that extracts the diagonal entries from a matrix (it should behave like `diag(x)` where `x` is a matrix).

**A5.** We can combine the existing functions to our advantage:

```{r}
x[!upper.tri(x) & !lower.tri(x)]

diag(x)
```

**Q6.** What does `df[is.na(df)] <- 0` do? How does it work?

**A6.** This command replaces every instance of `NA` in a dataframe with `0`.

`is.na(df)` produces a matrix of logical values, which provides a way to select and assign.

```{r}
(df <- tibble(x = c(1, 2, NA), y = c(NA, 5, NA)))

is.na(df)

class(is.na(df))
```



