# Vectors

```{r, include = FALSE}
source("common.R")
```

## Exercises 3.2.5 

**Q1.** How do you create raw and complex scalars? (See `?raw` and `?complex`.)

**A1.** In R, scalars are nothing but vectors of length 1, and can be created using the same constructor.

- Raw vectors

The raw type holds raw bytes, and can be created using `charToRaw()`. For example,

```{r}
x <- "A string"

(y <- charToRaw(x))

typeof(y)
```

An alternative is to use `as.raw()`:

```{r}
as.raw("–") # en-dash
as.raw("—") # em-dash
```

- Complex vectors

Complex vectors are used to represent (surprise!) complex numbers.

Example of a complex scalar:

```{r}
(x <- complex(length.out = 1, real = 1, imaginary = 8))

typeof(x)
```

**Q2.** Test your knowledge of the vector coercion rules by predicting the output of the following uses of `c()`:

```{r, eval=FALSE}
c(1, FALSE)
c("a", 1)
c(TRUE, 1L)
```

**A2.** The vector coercion rules dictate that the data type with smaller size will be converted to data type with bigger size.

```{r}
c(1, FALSE)

c("a", 1)

c(TRUE, 1L)
```

**Q3.** Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?

**A3.** The coercion rules for vectors reveal why some of these comparisons return the results that they do.

```{r}
1 == "1"

c(1, "1")
```

```{r}
-1 < FALSE

c(-1, FALSE)
```

```{r}
"one" < 2

c("one", 2)

sort(c("one", 2))
```

**Q4.** Why is the default missing value, `NA`, a logical vector? What's special about logical vectors? (Hint: think about `c(FALSE, NA_character_)`.)

**A4.** The `"logical"` type is the lowest in the coercion hierarchy.

So `NA` defaulting to any other type (e.g. `"numeric"`) would mean that any time there is a missing element in a vector, rest of the elements would be converted to a type higher in hierarchy, which would be problematic for types lower in hierarchy.

```{r}
typeof(NA)

c(FALSE, NA_character_)
```

**Q5.** Precisely what do `is.atomic()`, `is.numeric()`, and `is.vector()` test for?

**A5.** Let's discuss them one-by-one.

- `is.atomic()`

This function checks if the object is a vector of atomic *type* (or `NULL`).

Quoting docs:

> `is.atomic` is true for the atomic types ("logical", "integer", "numeric", "complex", "character" and "raw") and `NULL`.

```{r}
is.atomic(NULL)

is.atomic(list(NULL))
```

- `is.numeric()`

Its documentation says:

> `is.numeric` should only return true if the base type of the class is `double` or `integer` and values can reasonably be regarded as `numeric`

Therefore, this function only checks for `double` and `integer` base types and not other types based on top of these types (`factor`, `Date`, `POSIXt`, or `difftime`).

```{r}
is.numeric(1L)

is.numeric(factor(1L))
```

- `is.vector()`

As per its documentation:

> `is.vector` returns `TRUE` if `x` is a vector of the specified mode having no attributes *other than names*. It returns `FALSE` otherwise.

Thus, the function can be incorrectif the object has attributes other than `names`.

```{r}
x <- c("x" = 1, "y" = 2)

is.vector(x)

attr(x, "m") <- "abcdef"

is.vector(x)
```

A better way to check for a vector:

```{r}
is.null(dim(x))
```

## Exercises 3.3.4

**Q1.** How is `setNames()` implemented? How is `unname()` implemented? Read the source code.

**A1.** Let's have a look at implementations for these functions.

- `setNames()`

```{r}
setNames
```

Given this function signature, we can see why, when no first argument is given, the result is still a named vector.

```{r}
setNames(, c("a", "b"))

setNames(c(1, 2), c("a", "b"))
```

- `unname()`

```{r}
unname
```

`unname()` removes existing names (or dimnames) by setting them to `NULL`.

```{r}
unname(setNames(, c("a", "b")))
```

**Q2.** What does `dim()` return when applied to a 1-dimensional vector? When might you use `NROW()` or `NCOL()`?

**A2.** Dimensions for a 1-dimensional vector are `NULL`. For example,

```{r}
dim(c(1, 2))
```


`NROW()` and `NCOL()` are helpful for getting dimensions for 1D vectors by treating them as if they were matrices or dataframes.

```{r}
# example-1
x <- character(0)

dim(x)

nrow(x)
NROW(x)

ncol(x)
NCOL(x)

# example-2
y <- 1:4

dim(y)

nrow(y)
NROW(y)

ncol(y)
NCOL(y)
```

**Q3.** How would you describe the following three objects? What makes them different from `1:5`?

```{r}
x1 <- array(1:5, c(1, 1, 5))
x2 <- array(1:5, c(1, 5, 1))
x3 <- array(1:5, c(5, 1, 1))
```

**A3.** `x1`, `x2`, and `x3` are one-dimensional **array**s, but with different "orientations", if we were to mentally visualize them. 

`x1` has 5 entries in the third dimension, `x2` in the second dimension, while `x1` in the first dimension.

**Q4.** An early draft used this code to illustrate `structure()`:

```{r}
structure(1:5, comment = "my attribute")
```

But when you print that object you don't see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help.)

**A4.** From `?attributes` (emphasis mine):

> Note that some attributes (namely class, **comment**, dim, dimnames, names, row.names and tsp) are treated specially and have restrictions on the values which can be set.

```{r}
structure(1:5, x = "my attribute")

structure(1:5, comment = "my attribute")
```

## Exercises 3.4.5

**Q1.** What sort of object does `table()` return? What is its type? What attributes does it have? How does the dimensionality change as you tabulate more variables?

**A1.** `table()` returns an array of `integer` type and its dimensions scale with the number of variables present.

```{r}
(x <- table(mtcars$am))
(y <- table(mtcars$am, mtcars$cyl))
(z <- table(mtcars$am, mtcars$cyl, mtcars$vs))

# type
purrr::map(list(x, y, z), typeof)

# attributes
purrr::map(list(x, y, z), attributes)
```

**Q2.** What happens to a factor when you modify its levels? 

```{r, results = FALSE}
f1 <- factor(letters)
levels(f1) <- rev(levels(f1))
```

**A2.** Its levels change but the underlying integer values remain the same.

```{r}
f1 <- factor(letters)
f1
as.integer(f1)

levels(f1) <- rev(levels(f1))
f1
as.integer(f1)
```

**Q3.** What does this code do? How do `f2` and `f3` differ from `f1`?

```{r, results = FALSE}
f2 <- rev(factor(letters))
f3 <- factor(letters, levels = rev(letters))
```

**A3.** In this code:

- `f2`: Only the underlying integers are reversed, but levels remain unchanged.

```{r}
f2 <- rev(factor(letters))
f2
as.integer(f2)
```

- `f3`: Both the levels and the underlying integers are reversed.

```{r}
f3 <- factor(letters, levels = rev(letters))
f3
as.integer(f3)
```

## Exercises 3.5.4

**Q1.** List all the ways that a list differs from an atomic vector.

**A1.** Here is a table of comparison:

| feature                        | atomic vector                                    | list (aka generic vector)                              |
| :----------------------------- | :----------------------------------------------- | :----------------------------------------------------- |
| element type                   | unique                                           | mixed^[a list can contain a mix of types]              |
| recursive?                     | no                                               | yes^[a list can contain itself]                        |
| return for out-of-bounds index | `NA`                                             | `NULL`                                                 |
| memory address                 | single memory reference^[`lobstr::ref(c(1, 2))`] | reference per list element^[`lobstr::ref(list(1, 2))`] |

**Q2.** Why do you need to use `unlist()` to convert a list to an atomic vector? Why doesn't `as.vector()` work? 

**A2.** A list already *is* a (generic) vector, so `as.vector()` is not going to change anything, and there is no `as.atomic.vector`. Thus, we need to use `unlist()`.

```{r}
x <- list(a = 1, b = 2)

is.vector(x)
is.atomic(x)

# still a list
as.vector(x)

# now a vector
unlist(x)
```

**Q3.** Compare and contrast `c()` and `unlist()` when combining a date and date-time into a single vector.

**A3.** Let's first create a date and datetime object

```{r}
date <- as.Date("1947-08-15")
datetime <- as.POSIXct("1950-01-26 00:01", tz = "UTC")
```

And check their attributes and underlying `double` representation:

```{r}
attributes(date)
attributes(datetime)

as.double(date) # number of days since the Unix epoch 1970-01-01
as.double(datetime) # number of seconds since then
```

- Behavior with `c()`

Since `S3` method for `c()` dispatches on the first argument, the resulting class of the vector is going to be the same as the first argument. Because of this, some attributes will be lost.

```{r}
c(date, datetime)

attributes(c(date, datetime))

c(datetime, date)

attributes(c(datetime, date))
```

- Behavior with `unlist()`

It removes all attributes and we are left only with the underlying double representations of these objects.

```{r}
unlist(list(date, datetime))

unlist(list(datetime, date))
```

## Exercises 3.6.8

**Q1.** Can you have a data frame with zero rows? What about zero columns?

**A1.** Data frame with 0 rows is possible. This is basically a list with a vector of length 0.

```{r}
data.frame(x = numeric(0))
```

Data frame with 0 columns is also possible. This will be an empty list.

```{r}
data.frame(row.names = 1)
```

And, finally, data frame with 0 rows *and* columns is also possible:

```{r}
data.frame()

dim(data.frame())
```

Although, it might not be common to *create* such data frames, they can be results of subsetting. For example,

```{r}
BOD[0, ]

BOD[, 0]

BOD[0, 0]
```

**Q2.** What happens if you attempt to set rownames that are not unique?

**A2.** If you attempt to set data frame rownames that are not unique, it will not work.

```{r, error=TRUE}
data.frame(row.names = c(1, 1))
```

**Q3.** If `df` is a data frame, what can you say about `t(df)`, and `t(t(df))`? Perform some experiments, making sure to try different column types.

**A3.** Transposing a data frame:

- transforms it into a matrix 
- coerces all its elements to be of the same type

```{r}
# original
(df <- head(iris))

# transpose
t(df)

# transpose of a transpose
t(t(df))

# is it a dataframe?
is.data.frame(df)
is.data.frame(t(df))
is.data.frame(t(t(df)))

# check type
typeof(df)
typeof(t(df))
typeof(t(t(df)))

# check dimensions
dim(df)
dim(t(df))
dim(t(t(df)))
```

**Q4.** What does `as.matrix()` do when applied to a data frame with columns of different types? How does it differ from `data.matrix()`?

**A4.** The return type of `as.matrix()` depends on the data frame column types.

As docs for `as.matrix()` mention:

> The method for data frames will return a character matrix if there is only atomic columns and any non-(numeric/logical/complex) column, applying as.vector to factors and format to other non-character columns. Otherwise the usual coercion hierarchy (logical < integer < double < complex) will be used, e.g. all-logical data frames will be coerced to a logical matrix, mixed logical-integer will give an integer matrix, etc.

Let's experiment:

```{r}
# example with mixed types (coerced to character)
(df <- head(iris))

as.matrix(df)

str(as.matrix(df))

# another example (no such coercion)
BOD

as.matrix(BOD)
```

On the other hand, `data.matrix()` always returns a numeric matrix.

From documentation of `data.matrix()`:

> Return the matrix obtained by converting all the variables in a data frame to numeric mode and then binding them together as the columns of a matrix. Factors and ordered factors are replaced by their internal codes. [...] Character columns are first converted to factors and then to integers.

Let's experiment:

```{r}
data.matrix(df)

str(data.matrix(df))
```

## Session information

```{r}
sessioninfo::session_info(include_base = TRUE)
```
