# Vectors

```{r, include = FALSE}
source("common.R")
```

## Exercise 3.2.5 

**Q1.** Create raw and complex scalars

The raw type holds raw bytes. For example,

```{r}
x <- "A string"

(y <- charToRaw(x))

typeof(y)
```

You can use it to also figure out differences in similar characters:

```{r}
charToRaw("–") # en-dash
charToRaw("—") # em-dash
```

Complex vectors can be used to represent (surprise!) complex numbers.

Example of a complex scalar:

```{r}
(x <- complex(length.out = 1, real = 1, imaginary = 8))

typeof(x)
```

**Q2.** Vector coercion rules

Usually, the more *general* type would take precedence.

```{r}
c(1, FALSE)

c("a", 1)

c(TRUE, 1L)
```

Let's try some more examples.

```{r}
c(1.0, 1L)

c(1.0, "1.0")

c(TRUE, "1.0")
```

**Q3.** Comparisons between different types

The coercion in vectors reveal why some of these comparisons return the results that they do.

```{r}
1 == "1"

c(1, "1")
```

```{r}
-1 < FALSE

c(-1, FALSE)
```

```{r}
"one" < 2

c("one", 2)

sort(c("one", 2))
```

**Q4.** Why `NA` defaults to `"logical"` type

The `"logical"` type is the lowest in the coercion hierarchy.

So `NA` defaulting to any other type (e.g. `"numeric"`) would mean that any time there is a missing element in a vector, rest of the elements would be converted to a type higher in hierarchy, which would be problematic for types lower in hierarchy.

```{r}
typeof(NA)

c(FALSE, NA_character_)
```

**Q5.** Misleading variants of `is.*` functions

- `is.atomic()`

This functions checks if the object is of atomic *type* (or `NULL`), and not if it is an atomic *vector*.

Quoting docs:

> `is.atomic` is true for the atomic types ("logical", "integer", "numeric", "complex", "character" and "raw") and `NULL`.

```{r}
is.atomic(NULL)

is.vector(NULL)
```

- `is.numeric()`

Its documentation says:

> `is.numeric` should only return true if the base type of the class is `double` or `integer` and values can reasonably be regarded as `numeric`

Therefore, this function only checks for `double` and `integer` base types and not other types based on top of these types (`factor`, `Date`, `POSIXt`, or `difftime`).

```{r}
x <- factor(c(1L, 2L))

is.numeric(x)
```


- `is.vector()`

As the documentation for this function reveals:

> `is.vector` returns `TRUE` if `x` is a vector of the specified mode having no attributes *other than names*. It returns `FALSE` otherwise.

Thus, the function can be incorrect in presence if the object has attributes other than `names`.

```{r}
x <- c("x" = 1, "y" = 2)

is.vector(x)

attr(x, "m") <- "abcdef"

is.vector(x)
```

A better way to check for a vector:

```{r}
is.null(dim(x))
```

## Exercise 3.3.4

**Q1.** Reading source code

```{r}
setNames

setNames(c(1, 2), c("a", "b"))
```

```{r}
unname

A <- provideDimnames(N <- array(1:24, dim = 2:4))

unname(A, force = TRUE)
```

**Q2.** 1-dimensional vector

Dimensions for a 1-dimensional vector are `NULL`.

`NROW()` and `NCOL()` are helpful for getting dimensions for 1D vectors by treating them as if they were matrices or dataframes.

```{r}
x <- character(0)

dim(x)

nrow(x)
NROW(x)

ncol(x)
NCOL(x)
```

**Q3.** Difference between vectors and arrays

- `1:5` is a dimensionless **vector** 
- `x1`, `x2`, and `x3` are one-dimensional **array** 

```{r}
(x <- 1:5)
dim(x)

(x1 <- array(1:5, c(1, 1, 5)))
(x2 <- array(1:5, c(1, 5, 1)))
(x3 <- array(1:5, c(5, 1, 1)))

dim(x1)
dim(x2)
dim(x3)
```

We can look at the dim attribute

**Q4.** About `structure()`

From `?attributes` (emphasis mine):

> Note that some attributes (namely class, **comment**, dim, dimnames, names, row.names and tsp) are treated specially and have restrictions on the values which can be set.

```{r}
structure(1:5, x = "my attribute")

structure(1:5, comment = "my attribute")
```

## Exercise 3.4.5

**Q1.** `table()` function

`table()` returns an array with integer type and its dimensions scale with the number of variables present.

```{r}
(x <- table(mtcars$am))
(y <- table(mtcars$am, mtcars$cyl))
(z <- table(mtcars$am, mtcars$cyl, mtcars$vs))

# type
purrr::map(list(x, y, z), typeof)

# attributes
purrr::map(list(x, y, z), attributes)
```

**Q2.** Factor reversal

Its levels changes but the underlying integer values remain the same.

```{r}
f1 <- factor(letters)
f1
as.integer(f1)

levels(f1) <- rev(levels(f1))
f1
as.integer(f1)
```

**Q3.** Factor reversal-2

`f2`: Only the underlying integers are reversed, but levels remain unchanged.
`f3`: Both the levels and the underlying integers are reversed.

```{r}
f2 <- rev(factor(letters))
f2
as.integer(f2)

f3 <- factor(letters, levels = rev(letters))
f3
as.integer(f3)
```

## Exercise 3.5.4

**Q1.** Differences between list and atomic vector

feature | atomic vector  | list (aka generic vector)
------- | -------------- | --------------
element type | unique | mixed^[a list can contain a mix of types]
recursive? | no | yes^[(a list can contain itself)]
return for out-of-bounds index | `NA`^[(e.g. `c(1)[2]`)] | `NULL`^[(e.g. `list(1)[2]`)]
memory address | single memory reference^[`lobstr::ref(c(1, 2))`] | reference per list element^[`lobstr::ref(list(1, 2))`]


**Q2.** Converting a list to an atomic vector

List already *is* a vector, so `as.vector` is not going to change anything, and there is no `as.atomic.vector`. Thus the need to use `unlist()`.

```{r}
x <- list(a = 1, b = 2)

is.vector(x)
is.atomic(x)

as.vector(x)

unlist(x)
```

**Q3.** Comparing `c()` and `unlist()` for date and datetime

```{r}
# creating a date and datetime
date <- as.Date("1947-08-15")
datetime <- as.POSIXct("1950-01-26 00:01", tz = "UTC")

# check attributes
attributes(date)
attributes(datetime)

# check their underlying double representation
as.double(date) # number of days since the Unix epoch 1970-01-01
as.double(datetime) # number of seconds since then
```

Behavior with `c()`: Works as expected. Only odd thing is that it strips the `tzone` attribute.

```{r}
c(date, datetime)

attributes(c(date, datetime))

c(datetime, date)

attributes(c(datetime, date))
```

Behavior with `unlist()`: Removes all attributes and we are left only with the underlying double representations of these objects.

```{r}
unlist(list(date, datetime))

unlist(list(datetime, date))
```

## Exercise 3.6.8

**Q1.** Data frame with 0 dimensions

Data frame with 0 rows is possible. This is basically a list with a vector of length 0.

```{r}
data.frame(x = numeric(0))
```

Data frame with 0 columns is possible. This will be an empty list.

```{r}
data.frame(row.names = 1)
```

Both in one go:

```{r}
data.frame()

dim(data.frame())
```

**Q2.** Non-unique rownames

If you attempt to set rownames that are not unique, it will not work.

```{r, error=TRUE}
data.frame(row.names = c(1, 1))
```

**Q3.** Transposing dataframes

Transposing a dataframe transforms it into a matrix and coerces all its elements to be of the same type.

```{r}
# original
(df <- head(iris))

# transpose
t(df)

# transpose of a transpose
t(t(df))

# is it a dataframe?
is.data.frame(df)
is.data.frame(t(df))
is.data.frame(t(t(df)))

# check type
typeof(df)
typeof(t(df))
typeof(t(t(df)))

# check dimensions
dim(df)
dim(t(df))
dim(t(t(df)))
```

**Q4.** `as.matrix()` and dataframe

The return type of `as.matrix()` depends on dataframe column types.

```{r}
# example with mixed types (coerced to character)
(df <- head(iris))

as.matrix(df)

str(as.matrix(df))

# another example (no such coercion)
BOD

as.matrix(BOD)
```

From documentation of `data.matrix()`:

> Return the matrix obtained by converting all the variables in a data frame to numeric mode and then binding them together as the columns of a matrix.

So `data.matrix()` always returns a numeric matrix:

```{r}
data.matrix(df)

str(data.matrix(df))
```

